{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-22 thesis/education.tex
data-race

\item
{\bf Mutex.}
The Landslide-friendly {\tt mutex\_test} checks for the possibility of two contending threads
accessing a mutex-protected critical section simultaneously.
It includes one thread repeating once the lock, unlock cycle
\revisionminor{(i.e., one contender calls {\tt critical\_section()} twice)}
so as to check for unlock/lock races as well as lock/lock interactions,
and, as mentioned previously (\cref{sec:education-pebbles-tests}),
checks for data race access pairs inside the mutex implementation itself.
As students have free rein to design their mutex internals,
this
%refers to
\revisionminor{test probes}
the general class of mutex bugs
in which any number of things can go wrong, depending on the implementation,
leading to mutual exclusion (or otherwise assertion) failure.
Whether each student's submitted P2 solved or suffered from such bugs
was determined by checking their grade files for any mutex-related penalties (assessed by the TAs),
then me double-checking their implementations by hand to confirm.

@@

		which can arise when many data race candidates cause Quicksand to spawn many Landslide jobs,

@@

Though I did not ask this question on the CMU survey,
my experience handling student questions in person
suggests CMU students
\revisionminor{(as well as Professor Eckhardt)}
also mostly wish for
better explanations of data races and
more detail and clarity in the preemption traces.
Though I present the formal definition of data races in the lecture
(\cref{sec:education-pebbles-recruiting})
and refer back to it in Landslide's documentation and output,
showing a concrete example
in future iterations of the user guide
would go a long way to illustrate the abstract concepts.
The preemption traces could be improved by making it clearer that the stack trace in each cell of the table
represents executing the thread in question from wherever it previously left off (or its inception)
all the way until it reaches that stack trace, then preempting it to run another thread.
They could also easily report more diagnostic information;
for example, showing the sets of memory conflicts between each thread,
annotating the type of each preemption point (yield, mutex, or data race),
and/or indicating the adversarial memory access for each data race preemption point.

>>>>

\item
{\bf Mutex.}
The Landslide-friendly {\tt mutex\_test} checks for the possibility of two contending threads
accessing a mutex-protected critical section simultaneously.
It includes one thread repeating once the lock, unlock cycle
\revisionminor{(i.e., one contender calls {\tt critical\_section()} twice)}
so as to check for unlock/lock races as well as lock/lock interactions,
and, as mentioned previously (\cref{sec:education-pebbles-tests}),
checks for data-race access pairs inside the mutex implementation itself.
As students have free rein to design their mutex internals,
this
%refers to
\revisionminor{test probes}
the general class of mutex bugs
in which any number of things can go wrong, depending on the implementation,
leading to mutual exclusion (or otherwise assertion) failure.
Whether each student's submitted P2 solved or suffered from such bugs
was determined by checking their grade files for any mutex-related penalties (assessed by the TAs),
then me double-checking their implementations by hand to confirm.

@@

		which can arise when many data-race candidates cause Quicksand to spawn many Landslide jobs,

@@

Though I did not ask this question on the CMU survey,
my experience handling student questions in person
suggests CMU students
\revisionminor{(as well as Professor Eckhardt)}
also mostly wish for
better explanations of data races and
more detail and clarity in the preemption traces.
Though I present the formal definition of data races in the lecture
(\cref{sec:education-pebbles-recruiting})
and refer back to it in Landslide's documentation and output,
showing a concrete example
in future iterations of the user guide
would go a long way to illustrate the abstract concepts.
The preemption traces could be improved by making it clearer that the stack trace in each cell of the table
represents executing the thread in question from wherever it previously left off (or its inception)
all the way until it reaches that stack trace, then preempting it to run another thread.
They could also easily report more diagnostic information;
for example, showing the sets of memory conflicts between each thread,
annotating the type of each preemption point (yield, mutex, or data race),
and/or indicating the adversarial memory access for each data-race preemption point.

### 2018-12-22 thesis/intro.tex
data-race

\begin{enumerate}
	\item \revisionminor{{\bf Data-race preemption points and Iterative Deepening (\Cref{chap:quicksand}).}
		I will present Quicksand, a stateless model checking execution framework
		that manages multiple simultaneous Landslide instances
		to automatically cope with exponential explosion.
		Quicksand is powered by {\em Iterative Deepening},}
		a new algorithm for navigating the trade-off in how many preemption points to test at once.
		Iterative Deepening incorporates state space estimation \cite{estimation}
		to decide on-the-fly whether each state space is worth pursuing,
		and uses data race analysis \cite{tsan,fasttrack}
		to find new preemption point candidates based on a program's dynamic behaviour.
		\revision{This chapter will include a soundness proof,
		showing that testing only synchronization and data-race preemption points
		still constitutes a full formal verification of the test,
		and}
		a large evaluation, comparing its performance to three \revisionminor{prior-work} approaches
		across 600+ unique tests.
		I will show that \revisionminor{Quicksand}
		outperforms prior work in terms \revisionminor{of both} finding bugs quickly
		and verifying correctness when no bug exists.
	\item {\bf Educational use (\Cref{chap:education}).}
		For the past \revisionminor{seven} semesters,
		I have offered a fully-automated version of Landslide to students in 15-410,
		CMU's undergraduate Operating System Design and Implementation class,
		for use as a debugging aid during the thread library project \cite{thrlib}.
		%Recently
		I have also extended Landslide to handle Pintos kernel projects from other universities \cite{pintos}.
		In the two most recent semesters, I collaborated with Operating Systems course staff at the University of Chicago,
		\revisionminor{which uses Pintos,
		and at The Pennsylvania State University,
		which recently adopted CMU's thread library project,}
		to provide
		%debugging feedback to their students.
		\revisionminor{their students an opportunity to benefit from Landslide as well.}
		At all three universities I then collected statistics on the numbers and types of bugs found,
		and surveyed students to understand the human experience.
		This section will present the study's results
		\revisionminor{and prove that stateless model checking is suitable for use} in an educational setting.
	\item {\bf Hardware transactional memory (HTM) (\Cref{chap:tm}).}
		\revision{I will introduce a new concurrency model
		for stateless model checkers to emulate HTM's execution semantics
		in terms of existing concurrency primitives.
		This model is accompanied by a soundness proof
		which allows checkers to avoid simulating conflict abort rollbacks,
		reducing the state space substantially,
		while preserving the formal verification guarantee.
		I have implemented a new transactional testing mode in Landslide,
		with support for the many different abort reasons HTM introduces,
		optional weak atomicity semantics for simulating software TM instead,
		and a ``retry set'' reduction algorithm to identify and prune new types of equivalences.
		The evaluation includes several real-world HTM programs and benchmarks,
		and shows that Landslide can both find bugs and verify correctness
		using the new model with reasonable performance.}
\end{enumerate}

>>>>

\begin{enumerate}
	\item \revisionminor{{\bf Data-race preemption points and Iterative Deepening (\Cref{chap:quicksand}).}
		I will present Quicksand, a stateless model checking execution framework
		that manages multiple simultaneous Landslide instances
		to automatically cope with exponential explosion.
		Quicksand is powered by {\em Iterative Deepening},}
		a new algorithm for navigating the trade-off in how many preemption points to test at once.
		Iterative Deepening incorporates state space estimation \cite{estimation}
		to decide on-the-fly whether each state space is worth pursuing,
		and uses data-race analysis \cite{tsan,fasttrack}
		to find new preemption point candidates based on a program's dynamic behaviour.
		\revision{This chapter will include a soundness proof,
		showing that testing only synchronization and data-race preemption points
		still constitutes a full formal verification of the test,
		and}
		a large evaluation, comparing its performance to three \revisionminor{prior-work} approaches
		across 600+ unique tests.
		I will show that \revisionminor{Quicksand}
		outperforms prior work in terms \revisionminor{of both} finding bugs quickly
		and verifying correctness when no bug exists.
	\item {\bf Educational use (\Cref{chap:education}).}
		For the past \revisionminor{seven} semesters,
		I have offered a fully-automated version of Landslide to students in 15-410,
		CMU's undergraduate Operating System Design and Implementation class,
		for use as a debugging aid during the thread library project \cite{thrlib}.
		%Recently
		I have also extended Landslide to handle Pintos kernel projects from other universities \cite{pintos}.
		In the two most recent semesters, I collaborated with Operating Systems course staff at the University of Chicago,
		\revisionminor{which uses Pintos,
		and at The Pennsylvania State University,
		which recently adopted CMU's thread library project,}
		to provide
		%debugging feedback to their students.
		\revisionminor{their students an opportunity to benefit from Landslide as well.}
		At all three universities I then collected statistics on the numbers and types of bugs found,
		and surveyed students to understand the human experience.
		This section will present the study's results
		\revisionminor{and prove that stateless model checking is suitable for use} in an educational setting.
	\item {\bf Hardware transactional memory (HTM) (\Cref{chap:tm}).}
		\revision{I will introduce a new concurrency model
		for stateless model checkers to emulate HTM's execution semantics
		in terms of existing concurrency primitives.
		This model is accompanied by a soundness proof
		which allows checkers to avoid simulating conflict abort rollbacks,
		reducing the state space substantially,
		while preserving the formal verification guarantee.
		I have implemented a new transactional testing mode in Landslide,
		with support for the many different abort reasons HTM introduces,
		optional weak atomicity semantics for simulating software TM instead,
		and a ``retry set'' reduction algorithm to identify and prune new types of equivalences.
		The evaluation includes several real-world HTM programs and benchmarks,
		and shows that Landslide can both find bugs and verify correctness
		using the new model with reasonable performance.}
\end{enumerate}

### 2018-12-22 thesis/landslide.tex
write more blurbs for quicksand: eval, verif intros; expt bias; summary
data-race

\begin{itemize}
	\item {\tt within\_function FUNC} - adds {\tt FUNC} to an allowlist of functions required to appear in the current stack trace before identifying a preemption point (see \cref{sec:landslide-pps})
	\item {\tt without\_function FUNC} - as above, but a denylist instead of an allowlist
	\item {\tt within\_user\_function FUNC} - as two above but finds the function in the userspace test program rather than the kernel code.
	\item {\tt without\_user\_function FUNC} - difference to two above same as stated one above.
	\item {\tt data\_race ADDR TID LAST\_CALL CURRENT\_SYSCALL} - specifies a data-race preemption point.
		\begin{itemize}
			\llitem {\tt ADDR} shall be the code address (in hex) of the racing address,
			{\em before} the execution of which a preemption will be issued.
			\llitem {\tt TID} indicates a thread ID required to be running for this data race.
				To specify data race PPs across all threads at once, set {\tt FILTER\_DRS\_BY\_TID=0} (see next section).
			\llitem {\tt LAST\_CALL} indicates a code address required to be the site of the last {\tt call} instruction executed
				(similar to specifying a stack trace, but using a full stack trace here degrades performance too much),
				or 0 to not use this feature.
				From personal experience I found this option rather useless and recommend always supplying 0.
				For further discussion see \cref{sec:quicksand-pps}.
			\llitem {\tt CURRENT\_SYSCALL} indicates the system call number if a user-space data race comes from within a kernel system call which accesses user memory (Pebbles only).
				Usually 0 (i.e., not in kernel code) but {\tt deschedule}'s system call number is common as well.
		\end{itemize}
	\item {\tt input\_pipe FILENAME} - FIFO file used for receiving messages from Quicksand (e.g. to suspend or resume execution).
		Requires {\tt id\_magic} option to be set (next section below).
		The odds that a human user will find spiritual enlightenment through using this option by hand are infinitesimal.
	\item {\tt output\_pipe FILENAME} - as above but for sending messages.
\end{itemize}

@@

\item {\bf Memory analysis options}
\begin{itemize}
	\item {\tt PURE\_HAPPENS\_BEFORE=1} - select Pure Happens-Before (1) or Limited Happens-Before (2) (\cref{sec:background-hb});
		corresponds to {\tt -V}/{\tt -H} in \cref{sec:landslide-quicksand-options}.
	\item {\tt FILTER\_DRS\_BY\_TID=1} - configures whether to use the {\tt TID} parameter of {\tt data\_\allowbreak{}race} described above.
	\item {\tt FILTER\_DRS\_BY\_LAST\_CALL=0} - configures whether to use the {\tt LAST\_CALL} parameter of {\tt data\_race} described above.
	\item {\tt ALLOW\_LOCK\_HANDOFF=0} - configures lockset tracking to permit or disallow a lock taken by one thread to be released by another thread.%
		\footnote{If enabled, accesses performed by the second thread before unlocking will not be considered protected by that lock,
			as Landslide cannot infer what prior event abstractly represented the lock's ownership changing,
			leading to spurious data race reports.
			This could be solved in future work with a new annotation.}
	\item {\tt ALLOW\_REENTRANT\_MALLOC\_FREE=0} - allow two threads to be in {\tt malloc}, {\tt free}, or so on simultaneously without declaring it a bug.%
		\footnote{Used in Pintos, where those functions lock/unlock the heap mutex themselves rather than relying on a wrapper function to do so before invoking them.}
	\item {\tt TESTING\_MUTEXES=0} - configure ``mutex testing'' mode (1),
		in which the data race analysis will not consider a mutex's implementation to be protected by the mutex itself.
		In other words, the mutex's internal memory accesses will be flagged as data races,
		thereby enabling Landslide to verify the mutual exclusion property.
		Normally (0), Landslide assumes mutual exclusion is provided in order to efficiently find data races in the rest of the code.
		Quicksand will automatically set this option for P2s when {\tt -t mutex\_test} is specified.
\end{itemize}

@@

	\item {\tt ignore\_sym NAME SIZE} - specifies a global variable {\tt NAME} of a given {\tt SIZE} in bytes whose memory accesses should be ignored for the purposes of DPOR and data race analysis.

@@

	\item {\tt sched\_func NAME} - specifies a function whose memory accesses should all be ignored for the purposes of DPOR and data race analysis.

@@

	\item {\tt thrlib\_function NAME} - specifies a function whose memory accesses should be ignored both by the data race analysis and by DPOR.

@@

The following annotations enable Landslide to track locksets~\cite{eraser} for data race analysis.
If not provided, it will be as if Landslide assumes no guarantees about mutual exclusion or happens-before,
and hence will identify all memory conflicts as data races.
(Note that the corresponding instrumentation for P2s is achieved automatically,
as the names of the mutex interface are mandated by the project specification.)

\begin{itemize}
	\item {\tt tell\_landslide\_mutex\_locking(void *mutex\_addr)} - indicates the beginning of the lock routine for
		whatever synchronization API Landslide should treat as the primitive for data race detection.
		In Pintos this is the {\tt sema\_*()} function family; in Pebbles they may be called anything.
	\item {\tt tell\_landslide\_mutex\_blocking(int owner\_tid)} - called ``immediately'' before a thread becomes blocked on the mutex.
		Definition of ``immediately'' similar to the {\tt forking()} and friends annotations above.
		{\tt owner\_tid} allows Landslide to efficiently unblock/re-block threads when the mutex holder changes
		(rather than relying on heuristic yield-loop detection);
		see {\tt kern\_mutex\_block\_others()} and {\tt deadlocked()} in {\tt schedule.c} for implementation details.
	\item {\tt tell\_landslide\_mutex\_locking\_done(void *mutex\_addr)} - indicates the end of the lock routine.
	\item {\tt tell\_landslide\_mutex\_trylocking(void *mutex\_addr)} - indicates the beginning of the trylock routine (if present).
	\item {\tt tell\_landslide\_mutex\_trylocking\_done(void *mutex\_addr, int succeeded)} -
		indicates when a thread is finished trylocking, even if it failed to get the lock (indicated by {\tt succeeded}).
	\item {\tt tell\_landslide\_mutex\_unlocking(void *mutex\_addr)} - indicates the beginning of the unlock routine.
	\item {\tt tell\_landslide\_mutex\_unlocking\_done()} - indicates the end of the unlock routine.
\end{itemize}

@@

\section{Architecture}

@@

{\tt memory.c} is responsible for all manner of memory access analysis.
It tracks heap allocations, checks reads and writes in the heap region against same;
tracks reads and writes (in any region) from each thread
and checks them against each other
for DPOR (\cref{sec:landslide-dpor})
and data race analysis (\cref{sec:landslide-datarace}).
For userspace tests, it also tracks which virtual address space ({\tt cr3}) belongs to the test program
via a state machine of the test lifecycle,
which lets it avoid false positive heap errors from other programs which have differently-addressed heaps
({\tt check\_user\_address\_space()} and {\tt ignore\_user\_access()}.

@@

The {\tt data\_race} command specifies an instruction pointer value to identify as a data race preemption point.
It can optionally be qualified by a thread ID, most recent system call number, et cetera,
as described in \cref{sec:landslide-dynamicconfig},
and is queried through {\tt suspected\_data\_race()}.

@@

Whenever a conflict is identified,
\revisionminor{Landslide} also invokes the data race analysis (\cref{sec:landslide-datarace}).
It checks for {\tt free()}/access conflicts as well as access pairs,
effectively treating deallocation of a heap block as a ``poisoning'' write to its entire contents,
which is considered to conflict with accesses to any address therein
on the grounds that reordering may expose a use-after-free.

@@

\begin{itemize}
	\item Uses Dynamic Partial Order Reduction (DPOR) \cite{dpor}
		to prune equivalent interleavings while exploring the state space (\cref{sec:landslide-dpor}).
	\item Uses its simulator's memory tracing integration \cite{simics,bochs}
		to power data race analysis \cite{djit,fasttrack} (\cref{sec:landslide-datarace}),
		as well as DPOR.
		%and DPOR \cite{dpor} (\cref{sec:landslide-dpor}).
	\item Supports Preemption Sealing \cite{sealing} to configure the set of active preemption points
		and hence keep state space sizes manageable (\cref{sec:landslide-pps}).
	\item Extends DPOR with Recursive and Weighted Backtrack Estimation \cite{estimation}
		to predict overall ultimate state space size and completion time (\cref{sec:landslide-estimate}).
	\item Extends DPOR (at the user's option) with Iterative Context Bounding \cite{chess-icb}
		to heuristically uncover bugs faster should they exist,
		at the expense of overall completion time (\cref{sec:landslide-icb}).
\end{itemize}

@@

\begin{itemize}
	\item Extends Preemption Sealing for finer-grained control
		over whether individual functions and/or global variables
		should be considered for, or excluded from,
		DPOR and/or data-race analysis
		(\cref{sec:landslide-config-landslide};
		{\tt thrlib\_function} and {\tt ignore\_dr\_function}, respectively).
		This allows a test case author to precisely target the exact subset of code
		she is interested in testing,
		avoiding unnecessary state space inflation from trusted code.
	\item Adds a mutex testing mode to configure whether the
		data race analysis should consider locks' internal accesses
		(\cref{sec:landslide-dynamicconfig}; {\tt TESTING\_MUTEXES}).
		This allows the user to verify that her lock implementation
		correctly provides mutual exclusion
		(upcoming in \cref{sec:education-pebbles-tests}; {\tt mutex\_test}),
		before relying on them as trusted code in subsequent tests.
	\item Implements the sleep set extension to DPOR \cite{partial-order-methods,dpor,optimal-dpor},
		which identifies and skips further equivalences based on which memory accesses
		DPOR intended to reorder,
		%, rather than just switching threads and seeing what happens
		in a way that is slightly less optimal compared to prior work,
		but much simpler to implement
		and requires less state at runtime (\cref{sec:landslide-sleepsets}).
	\item Heuristically identifies infinite loops and livelocks
		by leveraging knowledge of how long the test program should be expected to run
		obtained during previous executions,
		uniquely suited to stateless model checking (\cref{sec:landslide-infloop}).
	\item Provides automatic code instrumentation facilities,
		beyond just the simple task of identifying important function boundaries,
		that attempt to understand a wide variety of even the most unusual student implementations
		(\cref{sec:landslide-pebblesspecifics},
		\cref{sec:landslide-pintosspecifics}).
		This includes an advanced heuristic for detecting ad-hoc synchronization loops
		that Landslide should treat as blocking, %a thread,
		even if it does not trigger the automatically-inserted blocking annotations
		(\cref{sec:landslide-blocking-yield}),
		and a further heuristic to ensure that any inaccuracy on the part of the first heuristic
		does not result in false-positive bug reports (\cref{sec:landslide-fp-deadlock}).
\end{itemize}

>>>>

\begin{itemize}
	\item {\tt within\_function FUNC} - adds {\tt FUNC} to an allowlist of functions required to appear in the current stack trace before identifying a preemption point (see \cref{sec:landslide-pps})
	\item {\tt without\_function FUNC} - as above, but a denylist instead of an allowlist
	\item {\tt within\_user\_function FUNC} - as two above but finds the function in the userspace test program rather than the kernel code.
	\item {\tt without\_user\_function FUNC} - difference to two above same as stated one above.
	\item {\tt data\_race ADDR TID LAST\_CALL CURRENT\_SYSCALL} - specifies a data-race preemption point.
		\begin{itemize}
			\llitem {\tt ADDR} shall be the code address (in hex) of the racing address,
			{\em before} the execution of which a preemption will be issued.
			\llitem {\tt TID} indicates a thread ID required to be running for this data race.
				To specify data-race preemption points across all threads at once, set {\tt FILTER\_DRS\_BY\_TID=0} (see next section).
			\llitem {\tt LAST\_CALL} indicates a code address required to be the site of the last {\tt call} instruction executed
				(similar to specifying a stack trace, but using a full stack trace here degrades performance too much),
				or 0 to not use this feature.
				From personal experience I found this option rather useless and recommend always supplying 0.
				For further discussion see \cref{sec:quicksand-pps}.
			\llitem {\tt CURRENT\_SYSCALL} indicates the system call number if a user-space data race comes from within a kernel system call which accesses user memory (Pebbles only).
				Usually 0 (i.e., not in kernel code) but {\tt deschedule}'s system call number is common as well.
		\end{itemize}
	\item {\tt input\_pipe FILENAME} - FIFO file used for receiving messages from Quicksand (e.g. to suspend or resume execution).
		Requires {\tt id\_magic} option to be set (next section below).
		The odds that a human user will find spiritual enlightenment through using this option by hand are infinitesimal.
	\item {\tt output\_pipe FILENAME} - as above but for sending messages.
\end{itemize}

@@

\item {\bf Memory analysis options}
\begin{itemize}
	\item {\tt PURE\_HAPPENS\_BEFORE=1} - select Pure Happens-Before (1) or Limited Happens-Before (2) (\cref{sec:background-hb});
		corresponds to {\tt -V}/{\tt -H} in \cref{sec:landslide-quicksand-options}.
	\item {\tt FILTER\_DRS\_BY\_TID=1} - configures whether to use the {\tt TID} parameter of {\tt data\_\allowbreak{}race} described above.
	\item {\tt FILTER\_DRS\_BY\_LAST\_CALL=0} - configures whether to use the {\tt LAST\_CALL} parameter of {\tt data\_race} described above.
	\item {\tt ALLOW\_LOCK\_HANDOFF=0} - configures lockset tracking to permit or disallow a lock taken by one thread to be released by another thread.%
		\footnote{If enabled, accesses performed by the second thread before unlocking will not be considered protected by that lock,
			as Landslide cannot infer what prior event abstractly represented the lock's ownership changing,
			leading to spurious data race reports.
			This could be solved in future work with a new annotation.}
	\item {\tt ALLOW\_REENTRANT\_MALLOC\_FREE=0} - allow two threads to be in {\tt malloc}, {\tt free}, or so on simultaneously without declaring it a bug.%
		\footnote{Used in Pintos, where those functions lock/unlock the heap mutex themselves rather than relying on a wrapper function to do so before invoking them.}
	\item {\tt TESTING\_MUTEXES=0} - configure ``mutex testing'' mode (1),
		in which the data-race analysis will not consider a mutex's implementation to be protected by the mutex itself.
		In other words, the mutex's internal memory accesses will be flagged as data races,
		thereby enabling Landslide to verify the mutual exclusion property.
		Normally (0), Landslide assumes mutual exclusion is provided in order to efficiently find data races in the rest of the code.
		Quicksand will automatically set this option for P2s when {\tt -t mutex\_test} is specified.
\end{itemize}

@@

	\item {\tt ignore\_sym NAME SIZE} - specifies a global variable {\tt NAME} of a given {\tt SIZE} in bytes whose memory accesses should be ignored for the purposes of DPOR and data-race analysis.

@@

	\item {\tt sched\_func NAME} - specifies a function whose memory accesses should all be ignored for the purposes of DPOR and data-race analysis.

@@

	\item {\tt thrlib\_function NAME} - specifies a function whose memory accesses should be ignored both by the data-race analysis and by DPOR.

@@

The following annotations enable Landslide to track locksets~\cite{eraser} for data-race analysis.
If not provided, it will be as if Landslide assumes no guarantees about mutual exclusion or happens-before,
and hence will identify all memory conflicts as data races.
(Note that the corresponding instrumentation for P2s is achieved automatically,
as the names of the mutex interface are mandated by the project specification.)

\begin{itemize}
	\item {\tt tell\_landslide\_mutex\_locking(void *mutex\_addr)} - indicates the beginning of the lock routine for
		whatever synchronization API Landslide should treat as the primitive for data-race detection.
		In Pintos this is the {\tt sema\_*()} function family; in Pebbles they may be called anything.
	\item {\tt tell\_landslide\_mutex\_blocking(int owner\_tid)} - called ``immediately'' before a thread becomes blocked on the mutex.
		Definition of ``immediately'' similar to the {\tt forking()} and friends annotations above.
		{\tt owner\_tid} allows Landslide to efficiently unblock/re-block threads when the mutex holder changes
		(rather than relying on heuristic yield-loop detection);
		see {\tt kern\_mutex\_block\_others()} and {\tt deadlocked()} in {\tt schedule.c} for implementation details.
	\item {\tt tell\_landslide\_mutex\_locking\_done(void *mutex\_addr)} - indicates the end of the lock routine.
	\item {\tt tell\_landslide\_mutex\_trylocking(void *mutex\_addr)} - indicates the beginning of the trylock routine (if present).
	\item {\tt tell\_landslide\_mutex\_trylocking\_done(void *mutex\_addr, int succeeded)} -
		indicates when a thread is finished trylocking, even if it failed to get the lock (indicated by {\tt succeeded}).
	\item {\tt tell\_landslide\_mutex\_unlocking(void *mutex\_addr)} - indicates the beginning of the unlock routine.
	\item {\tt tell\_landslide\_mutex\_unlocking\_done()} - indicates the end of the unlock routine.
\end{itemize}

@@

\section{Architecture}
\label{sec:landslide-architecture}

@@

{\tt memory.c} is responsible for all manner of memory access analysis.
It tracks heap allocations, checks reads and writes in the heap region against same;
tracks reads and writes (in any region) from each thread
and checks them against each other
for DPOR (\cref{sec:landslide-dpor})
and data-race analysis (\cref{sec:landslide-datarace}).
For userspace tests, it also tracks which virtual address space ({\tt cr3}) belongs to the test program
via a state machine of the test lifecycle,
which lets it avoid false positive heap errors from other programs which have differently-addressed heaps
({\tt check\_user\_address\_space()} and {\tt ignore\_user\_access()}.

@@

The {\tt data\_race} command specifies an instruction pointer value to identify as a data-race preemption point.
It can optionally be qualified by a thread ID, most recent system call number, et cetera,
as described in \cref{sec:landslide-dynamicconfig},
and is queried through {\tt suspected\_data\_race()}.

@@

Whenever a conflict is identified,
\revisionminor{Landslide} also invokes the data-race analysis (\cref{sec:landslide-datarace}).
It checks for {\tt free()}/access conflicts as well as access pairs,
effectively treating deallocation of a heap block as a ``poisoning'' write to its entire contents,
which is considered to conflict with accesses to any address therein
on the grounds that reordering may expose a use-after-free.

@@

\begin{itemize}
	\item Uses Dynamic Partial Order Reduction (DPOR) \cite{dpor}
		to prune equivalent interleavings while exploring the state space (\cref{sec:landslide-dpor}).
	\item Uses its simulator's memory tracing integration \cite{simics,bochs}
		to power data-race analysis \cite{djit,fasttrack} (\cref{sec:landslide-datarace}),
		as well as DPOR.
		%and DPOR \cite{dpor} (\cref{sec:landslide-dpor}).
	\item Supports Preemption Sealing \cite{sealing} to configure the set of active preemption points
		and hence keep state space sizes manageable (\cref{sec:landslide-pps}).
	\item Extends DPOR with Recursive and Weighted Backtrack Estimation \cite{estimation}
		to predict overall ultimate state space size and completion time (\cref{sec:landslide-estimate}).
	\item Extends DPOR (at the user's option) with Iterative Context Bounding \cite{chess-icb}
		to heuristically uncover bugs faster should they exist,
		at the expense of overall completion time (\cref{sec:landslide-icb}).
\end{itemize}

@@

\begin{itemize}
	\item Extends Preemption Sealing for finer-grained control
		over whether individual functions and/or global variables
		should be considered for, or excluded from,
		DPOR and/or data-race analysis
		(\cref{sec:landslide-config-landslide};
		{\tt thrlib\_function} and {\tt ignore\_dr\_function}, respectively).
		This allows a test case author to precisely target the exact subset of code
		she is interested in testing,
		avoiding unnecessary state space inflation from trusted code.
	\item Adds a mutex testing mode to configure whether the
		data-race analysis should consider locks' internal accesses
		(\cref{sec:landslide-dynamicconfig}; {\tt TESTING\_MUTEXES}).
		This allows the user to verify that her lock implementation
		correctly provides mutual exclusion
		(upcoming in \cref{sec:education-pebbles-tests}; {\tt mutex\_test}),
		before relying on them as trusted code in subsequent tests.
	\item Implements the sleep set extension to DPOR \cite{partial-order-methods,dpor,optimal-dpor},
		which identifies and skips further equivalences based on which memory accesses
		DPOR intended to reorder,
		%, rather than just switching threads and seeing what happens
		in a way that is slightly less optimal compared to prior work,
		but much simpler to implement
		and requires less state at runtime (\cref{sec:landslide-sleepsets}).
	\item Heuristically identifies infinite loops and livelocks
		by leveraging knowledge of how long the test program should be expected to run
		obtained during previous executions,
		uniquely suited to stateless model checking (\cref{sec:landslide-infloop}).
	\item Provides automatic code instrumentation facilities,
		beyond just the simple task of identifying important function boundaries,
		that attempt to understand a wide variety of even the most unusual student implementations
		(\cref{sec:landslide-pebblesspecifics},
		\cref{sec:landslide-pintosspecifics}).
		This includes an advanced heuristic for detecting ad-hoc synchronization loops
		that Landslide should treat as blocking, %a thread,
		even if it does not trigger the automatically-inserted blocking annotations
		(\cref{sec:landslide-blocking-yield}),
		and a further heuristic to ensure that any inaccuracy on the part of the first heuristic
		does not result in false-positive bug reports (\cref{sec:landslide-fp-deadlock}).
\end{itemize}

### 2018-12-22 thesis/overview.tex
data-race

A data race analysis may be either {\em static} (inspecting source code) \cite{racerx} or {\em dynamic} (tracking individual accesses arising at run-time) \cite{tsan}.
This paper focuses exclusively on dynamic analysis,
so although our example refers to numbered source lines for ease of explanation,
in practice we are actually classifying the individual memory access events corresponding to those lines during execution.
Actually, each {\tt x++} statement likely compiles to two separate load or store instructions, so each of those two instructions from each of the two marked source lines pairwise will race (except for the two loads, which are both reads).

@@

		The {\em heuristic} search ordering strategy and framework for incorporating {\em data race preemption points}

@@

		(e.g., applying {\em DPOR} without using {\em data race preemption points}).

>>>>

A data-race analysis may be either {\em static} (inspecting source code) \cite{racerx}
or {\em dynamic} (tracking individual accesses arising at run-time) \cite{tsan}.
This paper focuses exclusively on dynamic analysis,
so although our example refers to numbered source lines for ease of explanation,
in practice we are actually classifying the individual memory access events corresponding to those lines during execution.
Actually, each {\tt x++} statement likely compiles to two separate load or store instructions, so each of those two instructions from each of the two marked source lines pairwise will race (except for the two loads, which are both reads).

@@

		The {\em heuristic} search ordering strategy and framework for incorporating {\em data-race preemption points}

@@

		(e.g., applying {\em DPOR} without using {\em data-race preemption points}).
